Molecular dynamics simulation of synchronization in driven particles

### 2021-01-16 manuscript/manuscript_ajp_2020.tex
updated text around Fig.3

In Fig.~\ref{fig:0}(b) 
we show the $y-$position of the particle
as a function of time.
We 
normalize $y$ by $\lambda$, the spatial period of the potential
to show how the particle moves between substrate minima.  
The initial particle position is $y = 1 a_0$,
where the $F^l$ is negative.
%JUST BECAUSE
%causing the particle to move to
%a landscape minima at the start of the simulation.
As the driving force increases,
%the effects of the landscape become less obvious,
the motion of the particle
synchronizes with the driving period $T$.
The particle moves
in the positive y-direction
through $\Delta y = 2 \lambda$ over the time period $T$,
with 
the average velocity 
$\langle {v}_y \rangle= 2 \lambda f$. % where $T = 1/f$.
The horizontal dashed lines
coincide with the condition the substrate force is
minimum $F^l = -A_p$,  
i.e. $\sin{(\pi y / \lambda)} = 1$.
%or $\pi y / \lambda = m \pi/2$, where $m = 1, 5, 7$, {\it etc.}
%Once the particle motion
%coincides with the applied drive,
%indicating synchronization,
The %shows %indicates the
driving force is maximum when the landscape 
force on the particle is minimum,
as indicated with the coincidence of 
the particle position $y$
with the intersection of
the horizontal and vertical lines.
The slope of $y/\lambda$
is proportional to the net force on the particle.
This can be seen in 
Fig.~\ref{fig:0} %we demonstrate this relationship,
where local extrema appear in $y/\lambda$
when $F_{D}(t) \approx 0 $.

\begin{center}
\begin{figure}[h!]
\centering
\includegraphics[width=\columnwidth]{single_particle.pdf}
\caption{(a) The applied driving force $F^d(t)$. % (Eq.~\ref{eq:drive}).
  %where the parameter $F^{dc}$ (blue) is slowly increased from zero %to a constant value %as described in the text.
  (b) 
  The $y-$position of the driven particle
  as a function of time %of a single driven particle
  normalized by the period of the substrate $\lambda$.
  }
\label{fig:0}
\end{figure}
\end{center}

The competition between the driving force and landscape potential
can produce a variety of hopping patterns in the particle motion. 
%In Fig.~\ref{fig:0}
%reaching a minimum position
%of $y/\lambda = 1$
%over the total period $T = 100$.
%In Fig. ~\ref{fig:0}
%the average displacement during one time period is a single wavelength
%of the substrate $y(t_0+T) - y(t_0) = \lambda$.
%single dc case with figure
%dc_current 0.01
%ac_current 0.05
%ac_frequency 0.01
%fp_y 0.1
%
The relative values of $F^{ac}$, $F^{dc}$ and $A_p$
control the rate and distance a  particle moves 
forward and backward in the landscape.
To explore the possible hopping patterns,
we sweep through a range of $F^{dc}$ for fixed $F^{ac}$ and $A_p$,
as shown in Fig.~\ref{fig:1}(b).
We keep $F^{dc}$ constant
for $10^5$ simulation timesteps
and 
measure the average velocity $\langle v_y \rangle $ 
as a function of $F^{dc}$.
The average velocity $\langle{v}_y\rangle$
is the average displacement %$\Delta y = y(t_0+T) - y(t_0)$
over the period of the driving force.
We plot the results in Fig.~\ref{fig:1}
for $F^{ac} = 0.05$, $f=0.01$, and the landscape %force $\vec{F}^{l}$
shown in Fig.~\ref{fig:0}.
Discrete steps in $\langle v_y \rangle$ occur because 
the average velocity of the particle
is constant across a range of $F^{dc}$.
This synchronization
of the particle velocity can be measured
in terms of the landscape period.
%despite the overall magnitude
%of $F^{dc}$.
Each step represents a different pattern of hops
between substrate minima
performed by the particle.

If $A_p$ is large compared to the extrema of $F^{d}(t)$,
the particle oscillates back and forth
in a single minima
with no net velocity.
In Fig.~\ref{fig:1}, this occurs at 
low $F^{dc}$ when $\langle v_y \rangle$ is zero.

The particle 
achieves a variety of oscillation modes as a function of $F^{dc}$.
A mode is a periodic pattern of hops
with a constant average particle velocity, $\langle {v}_{y} \rangle$
over a range of driving forces $F^{dc}$.
We illustrate mode-locking in 
the velocity-force plot in Fig.~\ref{fig:1}.
Here $\langle v_{y} \rangle$ is increasing in non-uniform steps,
with a quantized height of
$\langle v_{y} \rangle = n \lambda f$,
where $n$ is an integer.  
%$\lambda = L/N_p = 36.5/20 = 1.825$ is the spatial period, or wavelength
%of the landscape,
%and $f = 0.01$ cycles per time unit.

Phase locked steps occur for average displacements integer
multiples of the substrate period $n\lambda$.

>>>>

In Fig.~\ref{fig:0}(b) 
we show the $y-$position of the particle
as a function of time.
We 
normalize $y$ by $\lambda$ %, the spatial period of the potential
to show how the particle moves between substrate minima.  
The initial particle position is $y = 0 $. %1 a_0$,
%where the $F^l$ is negative.
%JUST BECAUSE
%causing the particle to move to
%a landscape minima at the start of the simulation.
As the driving force increases,
%the effects of the landscape become less obvious,
the motion of the particle
synchronizes with the driving period $T$.
We measure 
average particle velocity $\langle{v}_y\rangle$
as the average displacement %$\Delta y = y(t_0+T) - y(t_0)$
over the period of the driving force.
The particle moves
in the positive y-direction
through $\Delta y = 2 \lambda$ over the time period $T$,
with 
the average velocity 
$\langle {v}_y \rangle= 2 \lambda f$. % where $T = 1/f$.
The horizontal dashed lines
coincide with the condition the substrate force is
minimum $F^l = -A_p$,  
i.e. $\sin{(\pi y / \lambda)} = 1$.
%or $\pi y / \lambda = m \pi/2$, where $m = 1, 5, 7$, {\it etc.}
%Once the particle motion
%coincides with the applied drive,
%indicating synchronization,
The %shows %indicates the
driving force is maximum when the landscape 
force on the particle is minimum,
as indicated with the coincidence of 
the particle position $y$
with the intersection of
the horizontal and vertical lines.
The slope of $y/\lambda$
is proportional to the net force on the particle.
This can be seen in 
Fig.~\ref{fig:0} %we demonstrate this relationship,
where local extrema appear in $y/\lambda$
when $F_{D}(t) \approx 0 $.

\begin{center}
\begin{figure}[h!]
\centering
\includegraphics[width=\columnwidth]{single_particle.pdf}
\caption{(a) The applied driving force $F^d(t)$. % (Eq.~\ref{eq:drive}).
  %where the parameter $F^{dc}$ (blue) is slowly increased from zero %to a constant value %as described in the text.
  (b) 
  The $y-$position of the driven particle
  %of a single driven particle
  normalized by the period of the substrate $\lambda$.
  %as a function of time
  }
\label{fig:0}
\end{figure}
\end{center}

The competition between the driving force and landscape potential
can produce a variety of hopping patterns in the particle motion. 
%In Fig.~\ref{fig:0}
%reaching a minimum position
%of $y/\lambda = 1$
%over the total period $T = 100$.
%In Fig. ~\ref{fig:0}
%the average displacement during one time period is a single wavelength
%of the substrate $y(t_0+T) - y(t_0) = \lambda$.
%single dc case with figure
%dc_current 0.01
%ac_current 0.05
%ac_frequency 0.01
%fp_y 0.1
%
The relative values of $F^{ac}$, $F^{dc}$ and $A_p$
control the rate and distance a  particle moves 
forward and backward in the landscape.
To explore the possible hopping patterns,
we sweep through a range of $F^{dc}$ for fixed $F^{ac}$ and $A_p$
to illustrate
a variety of oscillation modes as a function of $F^{dc}$.
A mode is a periodic pattern of hops
with a constant average particle velocity, $\langle {v}_{y} \rangle$
over a range of driving forces $F^{dc}$.
%as shown
In Fig.~\ref{fig:1} 
we keep $F^{dc}$ constant
for $10^5$ simulation timesteps
and 
measure the average velocity $\langle v_y \rangle $ 
as a function of $F^{dc}$,
with $F^{ac} = 0.05$, $f=0.01$, and the landscape %force $\vec{F}^{l}$
shown in Fig.~\ref{fig:0}.
%We plot the results in Fig.~\ref{fig:1} for 
The modes appear as 
discrete steps in $\langle v_y \rangle$ %occur because
across a range of $F^{dc}$,

%the average velocity of the particle
%is constant 
%despite the overall magnitude
%of $F^{dc}$.
Each step represents a different pattern of hops
between substrate minima
performed by the particle
due to the landscape confinement.  
In Fig.~\ref{fig:1} at %this occurs at 
low $F^{dc}$ the average velocity $\langle v_y \rangle$ is zero.
Since 
$A_p$ is large compared to the extrema of $F^{d}(t)$,
the particle oscillates back and forth
in a single minima with no net velocity.
At higher $F^{dc}$ the particle velocity 
$\langle v_{y} \rangle$ is increasing in steps of uniform height,
$\langle v_{y} \rangle = n \lambda f$,
where $n$ is an integer.
The step width is non-linear 
and can have a variety of
interesting patterns - including
a Devil's staircase pattern \cite{Bak1986}.
%$\lambda = L/N_p = 36.5/20 = 1.825$ is the spatial period, or wavelength
%of the landscape,
%and $f = 0.01$ cycles per time unit.

This synchronization
of the particle velocity can be measured
in terms of the landscape period.
Phase locked steps occur for average displacements integer
multiples of the substrate period $n\lambda$.
